---
title: "R Notebook"
output: html_notebook
---

# Settings and Library


```{r setup}
knitr::opts_knit$set(root.dir = '/home/francesco/Repository/Competition_of_Mechanisms/results')
```

```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(binom)
library(latex2exp)
library(tidyr)
library(ggpubr)
library(patchwork)


#set working directory for all the notebook knitr
print(getwd())
```

```{r}
palette <- c("GPT2" = "#003f5c", "GPT2-medium" = "#58508d", "GPT2-large" = "#bc5090", "GPT2-xl" = "#ff6361", "Pythia-6.9b" = "#ffa600")
FACTUAL_COLOR <- "#005CAB"
COUNTERFACTUAL_COLOR <- "#E31B23"
```

# GPT2 Small Configurations

```{r}
model <- "gpt2"
model_folder <- "gpt2_full"
n_layers <- 12
experiment <- "copyVSfact"
n_positions <- 12
positions_name <- c("-", "Subject", "2nd Subject", "3rd Subject", "Relation", "Relation Last", "Attribute*", "-", "Subject Repeat", "2nd Subject repeat", "3nd Subject repeat", "Relation repeat", "Last")
relevant_position <- c("Subject", "Relation", "Relation Last", "Attribute*", "Subject repeat", "Relation repeat", "Last")
example_position <- c("iPhone", "was developed", "by", "Google.", "iPhone", "was developed", "by")
n_relevant_position <- 7
layer_pattern <- c(11,10,10,10,9,9)
head_pattern <- c(10,0,7,10,6,9)
layer_pattern <- c(11,10,10,10,9,9)
head_pattern <- c(10,0,7,10,6,9)
factual_heads <- c(c(11,10),c(10,7))
factual_heads_layer <- c(11,10)
factual_heads_head <- c(10,7)
AXIS_TITLE_SIZE <- 60
AXIS_TEXT_SIZE <- 50
HEATMAP_SIZE <- 10
```

# Pythia-6.9b Small Configurations

```{r}
model <- "pythia-6.9b"
model_folder <- "pythia-6.9b_full"
n_layers <- 32
experiment <- "copyVSfact"
n_positions <- 12
positions_name <- c("-", "Subject", "2nd Subject", "3rd Subject", "Relation", "Relation Last", "Attribute*", "-", "Subject Repeat", "2nd Subject repeat", "3nd Subject repeat", "Relation repeat", "Last")
relevant_position <- c("Subject", "Relation", "Relation Last", "Attribute*", "Subject repeat", "Relation repeat", "Last")
n_relevant_position <- 7
layer_pattern <- c(10,10,15 ,17,17,19,19,20,20,21,23)
head_pattern <-  c(1,27, 17, 14,28,20,31,2, 18,8,25)
factual_heads <- c(c(11,10),c(10,7))
AXIS_TITLE_SIZE <- 60
AXIS_TEXT_SIZE <- 50
HEATMAP_SIZE <- 10
```

# Load functions:

```{r}

LOW_COLOR <- FACTUAL_COLOR
HIGH_COLOR <- COUNTERFACTUAL_COLOR

#LOW_COLOR <- "#1a80bb" # factual
#HIGH_COLOR <- "#a00000" # counterfactual
create_heatmap_base <- function(data, x, y, fill, midpoint = 0, text=FALSE) {
  # Convert strings to symbols for tidy evaluation
  x_sym <- rlang::sym(x)
  y_sym <- rlang::sym(y)
  fill_sym <- rlang::sym(fill)
  if (text==TRUE){
    p<- ggplot(data, aes(!!x_sym, !!y_sym, fill = !!fill_sym)) +
      geom_tile(colour = "grey") +
      scale_fill_gradient2(low = LOW_COLOR, mid = "white", high = HIGH_COLOR, midpoint = midpoint) +
      theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
      geom_text(aes(label = sprintf("%.2f", !!fill_sym)), color = "black", size = HEATMAP_SIZE)+
      labs(x = x, y = y)
  }else{
  p<- ggplot(data, aes(!!x_sym, !!y_sym, fill = !!fill_sym)) +
    geom_tile(colour = "grey") +
   scale_fill_gradient2(low = LOW_COLOR, mid = "white", high = HIGH_COLOR) +
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
   # geom_text(aes(label = sprintf("%.2f", !!fill_sym)), color = "black", size = HEATMAP_SIZE)+
    labs(x = x, y = y) 
  }
  return(p)
}
```

## Logit Lens - Residual Stream

### Functions

```{r}
create_heatmap <- function(data, x, y, legend, fill, high_color) {
  p <- create_heatmap_base(data, x, y, fill) +
    scale_fill_gradient2(low = "white", mid = "white", high = high_color, limits=c(-1,17), name = legend) +
    #scale_fill_gradient2(low = "black", mid = "white", high = high_color, name = "Logit") +
    #scale_fill_gradient2(low = "black", mid= "white", high = high_color, name = "Logit") +
    
    theme_minimal() +
    #addforce to have all the labels
    #scale_x_continuous(breaks = seq(0,n_layers-1 ,1)) +
    scale_y_continuous(breaks = seq(0,n_relevant_position -1,1), labels = relevant_position,)+
                      # sec.axis = dup_axis(name = "", labels =  example_position)) +
    scale_y_reverse(breaks = seq(0,n_relevant_position -1,1), labels = relevant_position, )+
                    #sec.axis = dup_axis(name = "", labels = example_position))+
    labs(x = "Layer", y = "")+
    #fix intenxity of fill
    theme(
      axis.text.x = element_text(size=AXIS_TEXT_SIZE),
      axis.text.y = element_text(size=AXIS_TEXT_SIZE,),
      #remove background grid
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title.x = element_text(size = AXIS_TITLE_SIZE),
      axis.title.y = element_text(size = AXIS_TITLE_SIZE),
      legend.text = element_text(size = 40),
      legend.title = element_text(size = AXIS_TEXT_SIZE),
      #remove the legend\
      legend.position = "bottom",
      #increase the legend size
      legend.key.size = unit(2.5, "cm"),
      # move the y ticks to the right
    ) 
  return(p)
}
```

```{r}
print(getwd())
```

### Load and process data

```{r}
#print current working directory
data <- read.csv(sprintf("%s/logit_lens/%s/logit_lens_data.csv", experiment, model_folder))
number_of_position <- max(as.numeric(data$position))
data_resid_post <- data %>% filter(grepl("resid_post", component))
data_resid_post$position_name <- positions_name[data_resid_post$position + 1]
#filter just the relevant positions
data_resid_post <- data_resid_post %>% filter(position == 1 | position==4 | position == 5 | position== 6 | position==8 |  position == 11 | position== 12)
unique_positions <- unique(data_resid_post$position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data_resid_post$mapped_position <- unname(position_mapping[as.character(data_resid_post$position)])
```

### Residual stram heatmaps

```{r}
p_fact <- create_heatmap(data_resid_post, "layer","mapped_position",bquote("Logit of " * italic(t[fact])), "mem",  FACTUAL_COLOR)
p_copy <- create_heatmap(data_resid_post, "layer", "mapped_position", bquote("Logit of " * italic(t[cofa])), "cp",  COUNTERFACTUAL_COLOR)
p_fact
p_copy
```

Save it:

```{r}
ggsave(sprintf("tmp/%s_%s_residual_stream/resid_post_mem.pdf", model, experiment), p_fact, width = 50, height = 32, units = "cm")
ggsave(sprintf("tmp/%s_%s_residual_stream/resid_post_cp.pdf", model, experiment), p_copy, width = 50, height = 32, units = "cm")
```

### Residual streams - lineplot index

```{r}
data_resid_post_altered <- data_resid_post %>% filter(position == 6)
data_resid_post_2_subject <- data_resid_post %>% filter(position == 8)
data_resid_post_last <- data_resid_post %>% filter(position ==12)
p_logit <-ggplot(data_resid_post_last, aes(x=layer))+
  #last
  geom_line(aes(y=mem, color="mem"),size=4,  alpha=0.8 )+
  geom_point(aes(y=mem, color="mem"),size=6, alpha=0.8)+
  geom_line(aes(y=cp, color="cp"),size=4,  alpha=0.8)+
  geom_point(aes(y=cp, color="cp"),size=6,  alpha=0.8)+
  scale_color_manual(values = c("mem" = FACTUAL_COLOR, "cp" = COUNTERFACTUAL_COLOR, "cp_alt"="darkblue", "mem_subj"="darkred"), labels=c("cp"= "Counterfactual Token","mem"="Factual Token", "cp_alt"= "Counterfactual Attribute","mem_subj"="Factual 2nd Subject")) +
  labs(y= "Logit in the Last Position", x="Layer", color="")+
  theme_minimal()+
 scale_x_continuous(breaks = seq(0,n_layers-1,1)) +
  scale_y_continuous(limits = c(0,17))+
 # scale_y_log10()+
  theme(
    axis.text.x = element_text(size=AXIS_TEXT_SIZE),
    axis.text.y = element_text(size=AXIS_TEXT_SIZE,),
    #remove background grid
    axis.title.x = element_text(size = AXIS_TITLE_SIZE),
    axis.title.y = element_text(size = AXIS_TITLE_SIZE),
    legend.text = element_text(size = AXIS_TEXT_SIZE),
    legend.title = element_text(size = AXIS_TEXT_SIZE),
    #remove the legend\
    legend.position = "bottom",
    #increase the legend size
    legend.key.size = unit(2, "cm"),
    aspect.ratio = 7/10,
    panel.border = element_rect(colour = "grey", fill=NA, size=1),
    # move the y ticks to the right
  ) + guides(color = guide_legend(ncol = 2, nrow=1))
p_logit
ggsave(sprintf("tmp/%s_%s_residual_stream/resid_post_all_linelogit_line.pdf", model, experiment), p_logit, width = 50, height = 30, units = "cm")
```

```{r}
```

save

```{r}
ggsave(sprintf("PaperPlot/%s_%s_residual_stream/resid_post_cp_index.pdf", model, experiment), p, width = 50, height = 30, units = "cm")
```

### Residual streams - lineplot logit

```{r}
data_resid_post_altered <- data_resid_post %>% filter(position == 6)
data_resid_post_2_subject <- data_resid_post %>% filter(position == 8)
data_resid_post_last <- data_resid_post %>% filter(position ==12)

p_idx<-ggplot(data_resid_post_altered, aes(x=layer))+
  geom_line(aes(y=mem_idx, color="mem"),size=4,  alpha=0.8 )+
  geom_point(aes(y=mem_idx, color="mem"),size=6, alpha=0.8)+
  geom_line(aes(y=cp_idx, color="cp"),size=4,  alpha=0.8)+
  geom_point(aes(y=cp_idx, color="cp"),size=6,  alpha=0.8)+
  scale_color_manual(values = c("mem" = FACTUAL_COLOR, "cp" = COUNTERFACTUAL_COLOR), labels=c( "Altered Token", "Factual Token")) +
  labs(y= bquote("Rank"), x="Layer", color="")+
  theme_minimal()+
  #scale_x_continuous(breaks = seq(0,n_layers,1)) +
  scale_y_log10()+
  theme(
    axis.text.x = element_text(size=AXIS_TEXT_SIZE),
    axis.text.y = element_text(size=AXIS_TEXT_SIZE,),
    #remove background grid
    axis.title.x = element_text(size = AXIS_TITLE_SIZE),
    axis.title.y = element_text(size = AXIS_TITLE_SIZE),
    legend.text = element_text(size = AXIS_TEXT_SIZE),
    legend.title = element_text(size = AXIS_TEXT_SIZE),
    #remove the legend\
    legend.position = "bottom",
    #increase the legend size
    legend.key.size = unit(2, "cm"),
    # move the y ticks to the right
  )
p_idx
```
```{r}
model
```
Save

```{r}
ggsave(sprintf("tmp/%s_%s_residual_stream/resid_post_index.pdf", model, experiment), p_idx, width = 50, height = 30, units = "cm")
```

### Multiple plots
#### With logit
```{r}
spacer <- plot_spacer()
p_logit <- p_logit + theme(aspect.ratio = 7/10, legend.position = c(0.5,-0.2))
p <- (p_fact / p_copy) | spacer | p_logit
p <- p + plot_layout(widths = c(0.8, 0.4, 1.3))
ggsave(sprintf("tmp/%s_%s_residual_stream/resid_post_all_linelogit.pdf", model, experiment), p, width = 100, height = 50, units = "cm")
#ggsave("resid_post_all_linelogit_new.pdf", p, width = 100, height = 50, units = "cm")
```
#### With index
```{r}
spacer <- plot_spacer()
p_idx <- p_idx + theme(aspect.ratio = 7/10, legend.position = c(0.5,-0.2))
p <- (p_fact / p_copy) | spacer | p_idx
p <- p + plot_layout(widths = c(0.8, 0.3, 1.3))
ggsave(sprintf("PaperPlot/%s_%s_residual_stream/resid_post_all_lineindex.pdf", model, experiment), p, width = 100, height = 50, units = "cm")

```
## Logit Attribution
### Functions
```{r}
create_heatmap <- function(data, x, y, fill, head=FALSE) {
  if(head){
    scale_x <- scale_x_discrete(breaks = seq(0,n_layers,1)) 
    angle = 0
  } else {
    scale_x <- scale_x_discrete(breaks = seq(0, n_positions,1), labels = positions_name)
    angle = 90
  }
  print(n_positions)
  p <- create_heatmap_base(data, x, y, fill) +
    theme_minimal() +
    #addforce to have all the labels
    scale_y_discrete(breaks = seq(0,n_layers,1)) +
    scale_x +
    labs(fill = "Logit Diff") +
    theme(
      axis.text.x = element_text(size=AXIS_TEXT_SIZE, angle = angle),
      axis.text.y = element_text(size=AXIS_TEXT_SIZE),
      #remove background grid
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title.x = element_text(size = AXIS_TITLE_SIZE),
      axis.title.y = element_text(size = AXIS_TITLE_SIZE),
      legend.text = element_text(size = 30),
      legend.title = element_text(size = 30),
      #remove the legend\
      legend.position = "bottom",
      # increase the size of the legend
      legend.key.size = unit(2, "cm"),
      # move the y ticks to the right
    )
  return(p)
}
```
### Load and process data
```{r}
data <- read.csv(sprintf("%s/logit_attribution/%s/logit_attribution_data.csv", experiment, model_folder))
```
### Head
```{r}
data_head <- data %>% filter(grepl("^L[0-9]+H[0-9]+$", label))
number_of_position <- max(as.numeric(data_head$position))
## filter to have just position 12
data_head_ <- data_head %>% filter(position == number_of_position)
# for each row split L and H and create a new column for each
data_head_ <- data_head_ %>% separate(label, c("layer", "head"), sep = "H")
#renominating the columns layer and head to Layer and Head
#remove L from layer
data_head_$layer <- gsub("L", "", data_head_$layer)

max_layer <- max(as.numeric(data_head_$layer))
max_head <- max(as.numeric(data_head_$head))
data_head_$layer <- factor(data_head_$layer, levels = c(0:max_layer))
data_head_$head <- factor(data_head_$head, levels = c(0:max_head))
colnames(data_head_)[1] <- "Layer"
colnames(data_head_)[2] <- "Head"
data_head_$diff_mean <- -data_head_$diff_mean
```

```{r}
p <- create_heatmap_base(data_head_, "Layer", "Head", "diff_mean") +
  theme_minimal() +
  #addforce to have all the labels
  scale_y_discrete(breaks = seq(0,n_layers,1)) +
  scale_x_discrete(breaks = seq(0,n_layers,1))  +
  labs(fill = expression(Delta[cofa])) +
  theme(
    axis.text.x = element_text(size=AXIS_TEXT_SIZE -8, angle = 0),
    axis.text.y = element_text(size=AXIS_TEXT_SIZE-8),
    #remove background grid
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = AXIS_TITLE_SIZE),
    axis.title.y = element_text(size = AXIS_TITLE_SIZE),
    legend.text = element_text(size = 40),
    legend.title = element_text(size = 70),
    #remove the legend\
    legend.position = "bottom",
    # increase the size of the legend
    legend.key.size = unit(2.5, "cm"),
    # move the y ticks to the right
  )
ggsave(sprintf("tmp/%s_%s_logit_attribution/logit_attribution_head_position%s.pdf", model, experiment, number_of_position), p, width = 60, height = 60, units = "cm")

```

Then perform some computation on the heads:
```{r}
### count the impact of the positive head ###
#sum all the negative values
factual_impact <- data_head_ %>% group_by(Layer) %>% summarise(positive_impact = sum(diff_mean[diff_mean < 0]))
# sum also across layers
factual_impact <- factual_impact %>% summarise(positive_impact = sum(positive_impact))
l10h7 <- data_head_ %>% filter(Layer == 10, Head == 7)
l10h7 <- l10h7$diff_mean
l11h10 <- data_head_ %>% filter(Layer == 11, Head == 10)
l11h10 <- l11h10$diff_mean

l10h7 <- 100 * l10h7 / sum(factual_impact)
print(l10h7)
l11h10 <- 100 * l11h10 / sum(factual_impact)
print(l11h10)

```

### MLP and Attention Barplot
```{r}
data_mlp <- data %>% filter(grepl("^[0-9]+_mlp_out$", label))
data_mlp <- data_mlp %>% separate(label, c("layer"), sep = "_mlp_out")
max_position <- max(as.numeric(data_mlp$position))
data_mlp <- data_mlp %>% filter(position == max_position)
data_attn <- data %>% filter(grepl("^[0-9]+_attn_out$", label))
data_attn <- data_attn %>% separate(label, c("layer"), sep = "_attn_out")
max_position <- max(as.numeric(data_mlp$position))
data_attn <- data_attn %>% filter(position == max_position)
#merge the two dataframe
data_barplot <- data_mlp
data_barplot$attn_dif <- data_attn$diff_mean
data_barplot$attc_cp <- data_attn$cp_mean
data_barplot$attc_mem <- data_attn$mem_mean
#rename columns diff_mean to mlp_dif
data_barplot <- data_barplot %>% rename("mlp_dif" = diff_mean)
data_barplot <- data_barplot %>% rename("mlp_cp" = cp_mean)
data_barplot <- data_barplot %>% rename("mlp_mem" = mem_mean)

#pivoting attn in order to plot mem and cp in the same barplot
data_attn <- data_barplot %>% pivot_longer(cols = c("attc_cp", "attc_mem"), names_to = "Block", values_to = "value")
data_mlp <- data_barplot %>% pivot_longer(cols = c("mlp_cp", "mlp_mem"), names_to = "Block", values_to = "value")
#modify mlp_cp and mlp_mem to Altered and Factual
data_mlp$Block <- gsub("mlp_", "", data_mlp$Block)
data_mlp$Block <- gsub("cp", "Altered", data_mlp$Block)
data_mlp$Block <- gsub("mem", "Factual", data_mlp$Block)
data_attn$Block <- gsub("attc_", "", data_attn$Block)
data_attn$Block <- gsub("cp", "Altered", data_attn$Block)
data_attn$Block <- gsub("mem", "Factual", data_attn$Block)
data_barplot <- data_barplot %>% rename("MLP Block" = mlp_dif )
data_barplot <- data_barplot %>% rename("Attention Block" = attn_dif)


data_barplot$`MLP Block` <- -data_barplot$`MLP Block`
data_barplot$`Attention Block` <- -data_barplot$`Attention Block`

data_barplot$layer <- as.numeric(data_barplot$layer) 
```
#### MLP
```{r}
ggplot(data_barplot, aes(x = as.numeric(layer), y = `MLP Block`, fill = "MLP Block")) +
  geom_col(position = position_dodge(), color="black", size=1) +
  labs(x = "Layer", y = expression(Delta[cofa]), fill = "") + # Naming the legend
  theme_minimal() +
  scale_fill_manual(values = c("MLP Block" = "#bc5090")) + # Assigning color to the "MLP Block"
  #scale_y_continuous(limits = c(-1, 1.5)) +
  scale_x_continuous(breaks= seq(0, n_layers-1, 1), labels = as.character(seq(0,n_layers-1,1))) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(size = 60),
    axis.text.y = element_text(size = 60),
    axis.title.x = element_text(size = 60),
    axis.title.y = element_text(size = 70),
    legend.text = element_text(size = 50),
    legend.title = element_text(size = 55),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2.5)) # Adjusting the legend
ggsave("mlp_block_norm.pdf", width = 50, height = 30, units = "cm")
#ggsave(sprintf("PaperPlot/%s_%s_logit_attribution/logit_mlp_position%s_diff.pdf", model, experiment, max_position), width = 50, height = 30, units = "cm")

```

#### Attention Out
```{r}
ggplot(data_barplot, aes(x = as.numeric(layer), y = `Attention Block`, fill = "Attention Block")) +
  geom_col(position = position_dodge(), color="black",size=1) +
  labs(x = "Layer", y = expression(Delta[cofa]), fill = "") + # Naming the legend
  theme_minimal() +
  scale_fill_manual(values = c("Attention Block" = "#ffa600")) + # Assigning color to the "MLP Block"
  #scale_y_continuous(limits = c(-1, 1.5)) +
  scale_x_continuous(breaks= seq(0, n_layers-1, 1), labels = as.character(seq(0,n_layers-1,1))) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(size = 60),
    axis.text.y = element_text(size = 60),
    axis.title.x = element_text(size = 60),
    axis.title.y = element_text(size = 70),
    legend.text = element_text(size = 50),
    legend.title = element_text(size = 55),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2.5)) # Adjusting the legend
ggsave("attn_block_norm.pdf", width = 50, height = 30, units = "cm")

#ggsave(sprintf("PaperPlot/%s_%s_logit_attribution/logit_attn_position%s_diff.pdf", model, experiment, max_position), width = 50, height = 30, , units = "cm")

```

#### HeatMaps
##### MLP
```{r}
data_mlp <- data %>% filter(grepl("^[0-9]+_mlp_out$", label))
data_mlp <- data_mlp %>% separate(label, c("layer"), sep = "_mlp_out")
max_layer <- max(as.numeric(data_mlp$layer))
max_position <- max(as.numeric(data_mlp$position))
#create layer column

data_mlp$layer <- factor(data_mlp$layer, levels = c(0:max_layer))
data_mlp$position <- factor(data_mlp$position, levels = c(0:max_position))

colnames(data_mlp)[1] <- "Layer"
colnames(data_mlp)[2] <- "Position"

data_mlp <- data_mlp %>% filter(Position == 1 | Position==4 | Position == 5 | Position== 6 | Position==8 |  Position == 11 | Position== 12)
unique_positions <- unique(data_mlp$Position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data_mlp$mapped_position <- unname(position_mapping[as.character(data_mlp$Position)])
data_mlp$Layer <- as.numeric(data_mlp$Layer) +1

relevant_position <- c("Subject", "Relation", "Relation Last", "Attribute*", "Subject repeat", "Relation repeat", "Last")
n_relevant_position <- 7
data_mlp$diff_mean <- -data_mlp$diff_mean
p <- create_heatmap_base(data_mlp, "Layer", "mapped_position", "diff_mean") +
  theme_minimal() +
  #addforce to have all the labels
  scale_x_continuous(breaks = seq(1,max_layer+1,1), labels=as.character(seq(1,max_layer+1,1))) +
  scale_y_reverse(breaks = seq(0,n_relevant_position-1, 1), labels=relevant_position) +
  labs(fill = "Logit Diff", y="") +
  theme(
    axis.text.x = element_text(size=AXIS_TEXT_SIZE),
    axis.text.y = element_text(size=AXIS_TEXT_SIZE),
    #remove background grid
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = AXIS_TITLE_SIZE),
    axis.title.y = element_text(size = AXIS_TITLE_SIZE),
    legend.text = element_text(size = 35),
    legend.title = element_text(size = AXIS_TITLE_SIZE),
    #remove the legend\
    legend.position = "bottom",
    # increase the size of the legend
    legend.key.size = unit(3, "cm"),
    # move the y ticks to the right
  )

p
ggsave(sprintf("PaperPlot/%s_%s_logit_attribution/logit_attribution_mlp_out.pdf", model, experiment), p, width = 50, height = 50, units = "cm")


```

##### Attn out
```{r}
#filter position f"{i}_mlp_out"
data_attn <- data %>% filter(grepl("^[0-9]+_attn_out$", label))
data_attn <- data_attn %>% separate(label, c("layer"), sep = "_attn_out")
max_layer <- max(as.numeric(data_attn$layer))
max_position <- max(as.numeric(data_attn$position))
#create layer column

data_attn$layer <- factor(data_attn$layer, levels = c(0:max_layer))
data_attn$position <- factor(data_attn$position, levels = c(0:max_position))

colnames(data_attn)[1] <- "Layer"
colnames(data_attn)[2] <- "Position"
data_attn <- data_attn %>% filter(Position == 1 | Position==4 | Position == 5 | Position== 6 | Position==8 |  Position == 11 | Position== 12)
unique_positions <- unique(data_attn$Position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data_attn$mapped_position <- unname(position_mapping[as.character(data_attn$Position)])
data_attn$Layer <- as.numeric(data_attn$Layer) -1

relevant_position <- c("Subject", "Relation", "Relation Last", "Attribute*", "Subject repeat", "Relation repeat", "Last")
n_relevant_position <- 7
data_attn$diff_mean <- -data_attn$diff_mean

p <- create_heatmap_base(data_attn, "Layer", "mapped_position", "diff_mean") +
  theme_minimal() +
  #addforce to have all the labels
  scale_x_continuous(breaks = seq(1,max_layer+1,1), labels=as.character(seq(1,max_layer+1,1))) +
  scale_y_reverse(breaks = seq(0,n_relevant_position-1, 1), labels=relevant_position) +
  labs(fill = "Logit Diff", y="") +
  theme(
    axis.text.x = element_text(size=AXIS_TEXT_SIZE),
    axis.text.y = element_text(size=AXIS_TEXT_SIZE),
    #remove background grid
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = AXIS_TITLE_SIZE),
    axis.title.y = element_text(size = AXIS_TITLE_SIZE),
    legend.text = element_text(size = 35),
    legend.title = element_text(size = AXIS_TITLE_SIZE),
    #remove the legend\
    legend.position = "bottom",
    # increase the size of the legend
    legend.key.size = unit(3, "cm"),
    # move the y ticks to the right
  )

p
ggsave(sprintf("PaperPlot/%s_%s_logit_attribution/logit_attribution_attn_out.pdf", model, experiment), p, width = 50, height = 50, units = "cm")


```

## Head Pattern
```{r}
data <- read.csv(sprintf("%s/head_pattern/%s/head_pattern_data.csv", experiment, model_folder))
```
### Last Position
```{r}

data_filtered <- data %>% filter(source_position == 12)
pattern_df <- data.frame(layer = layer_pattern, head = head_pattern)

data_final <- data_filtered %>% 
  inner_join(pattern_df, by = c("layer", "head"))
# Step 3: Prepare the data for plotting
data_final$y_label <- paste("Layer ", data_final$layer, " | Head ", data_final$head, sep="")
#filter just the relevant positions
data_final <- data_final %>% filter(dest_position == 1 | dest_position==4 | dest_position == 5 | dest_position== 6 | dest_position==8 |  dest_position == 11 | dest_position== 12)
unique_positions <- unique(data_final$dest_position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data_final$mapped_position <- unname(position_mapping[as.character(data_final$dest_position)])
# Create and plot the heatmap
data_final <- data_final %>%
  mutate(color = ifelse((
    y_label =="Layer 10 | Head 27" | 
    y_label=="Layer 17 | Head 28" |
    y_label=="Layer 20 | Head 18" |
    y_label=="Layer 21 | Head 8"
    ), "Target", "Other")) # Add color column
#gpt2 
# data_final <- data_final %>%
#   mutate(color = ifelse((y_label =="Layer 10 | Head 7" | y_label=="Layer 11 | Head 10"), "Target", "Other")) # Add color column
#pythia
data_final <- data_final %>%
 mutate(color = ifelse((y_label =="Layer 21 | Head 8" | y_label=="Layer 20 | Head 18" | y_label=="Layer 17 | Head 28" |  y_label=="Layer 10 | Head 27"), "Target", "Other")) # Add color column

library(ggnewscale) # for using new color scales within the same plot
# Your original plot for 'Other'
```

```{r}

max_positions <- max(data_final$mapped_position)
tile_width <- 1 
tile_height <- 1 


heatmap_plot <- ggplot(data_final %>% filter(color == "Target"), aes(x = mapped_position, y = y_label, fill = value)) +
  geom_tile(colour = "grey", width = tile_width, height = tile_height) +
  scale_x_continuous(breaks = seq(0, length(relevant_position) - 1,1), labels = relevant_position) +
  scale_y_discrete(limits = unique(data_final$y_label)) +
  scale_fill_gradient(low = "white", high = FACTUAL_COLOR, limits=range(data_final$value)) +
  labs(fill = "Attention\nScore:") +
  theme_minimal() +
  new_scale_fill() + # This tells ggplot to start a new fill scale
  geom_tile(data = data_final %>% filter(color == "Other"), aes(x = mapped_position, y = y_label, fill = value), width = tile_width, height = tile_height, colour="grey") +
  scale_fill_gradient(low = "white", high = COUNTERFACTUAL_COLOR, limits=range(data_final$value)) +
  scale_x_continuous(breaks = seq(0, length(relevant_position) - 1,1), labels = relevant_position) +
  scale_y_discrete(limits = unique(data_final$y_label)) +
  labs(fill = "Attention\nScore:") +
  theme(
    axis.text.x = element_text(size=60, angle = 45, hjust = 1),
    axis.text.y = element_text(size=60, angle = 0),
   # axis.text.x = element_text(size=40, angle = 45, hjust = 1),
   # axis.text.y = element_text(size=40, angle = 0),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.text = element_text(size = 45),
    legend.title = element_text(size = 50),
    legend.position = "right",
    legend.key.size = unit(1.3, "cm"),
  )
heatmap_plot
ggsave(sprintf("tmp/%s_%s_heads_pattern/head_pattern_layer.pdf", model, experiment), heatmap_plot, width = 53, height = 38, units = "cm")

```



### Full position
#### FUnctions
```{r}
create_heatmap <- function(data, x, y, fill,title, color) {
    p <- create_heatmap_base(data, x, y, fill) +
      scale_fill_gradient2(low = LOW_COLOR, mid = "white", high = color, midpoint = 0, limits=c(0,0.45)) +
      theme_minimal() +
      #addforce to have all the labels
      scale_y_discrete(breaks = seq(0, length(relevant_position)-1,1), labels = relevant_position) +
      scale_x_discrete(breaks = seq(0, length(relevant_position)-1,1), labels = relevant_position) +
      labs(fill = "Attention\nscore:", title=title) +
      theme(
        axis.text.x = element_text(size=AXIS_TEXT_SIZE-10, angle = 45, hjust = 1),
        axis.text.y = element_text(size=AXIS_TEXT_SIZE-10),
        title = element_text(size = AXIS_TEXT_SIZE-10),
        #remove background grid
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 30),
        legend.title = element_text(size = 40),
        #remove the legend\
        legend.position = "right",
        legend.key.size = unit(1.3, "cm"),
        # move the y ticks to the right
      )
    return(p)
}
  plot_pattern <- function(l, h, data,color){
    selected_layer <- l
    selected_head <- h
    data_head <- data %>% filter(layer == selected_layer & head == selected_head)
    max_source_position <- max(as.numeric(data_head$source_position))
    max_dest_position <- max(as.numeric(data_head$dest_position))
    data_head$source_position <- factor(data_head$source_position, levels = c(0:max_source_position))
    data_head$dest_position <- factor(data_head$dest_position, levels = c(0:max_dest_position))
    #filter just the relevant positions for both source_position and dest_position
    data_head <- data_head %>% filter(dest_position == 1 | dest_position==4 | dest_position == 5 | dest_position== 6 | dest_position==8 |  dest_position == 11 | dest_position== 12)
    data_head <- data_head %>% filter(source_position == 1 | source_position==4 | source_position == 5 | source_position== 6 | source_position==8 |  source_position == 11 | source_position== 12)
    #remap the position
    unique_positions <- unique(data_head$dest_position)
    position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
    # Apply the mapping to create a new column
    data_head$dest_mapped <- unname(position_mapping[as.character(data_head$dest_position)])
    data_head$source_mapped <- unname(position_mapping[as.character(data_head$source_position)])
    # order the position
    data_head$dest_mapped <- factor(data_head$dest_mapped, levels = c(0,1,2,3,4,5,6))
    data_head$source_mapped <- factor(data_head$source_mapped, levels = c(6,5,4,3,2,1,0))
    
    #reorder the source_position and dest_position contrary to the order of the factor
    
    #data_head$source_mapped <- factor(data_head$source_mapped, levels = rev(levels(data_head$source_mapped)))

    p <- create_heatmap(data_head, "dest_mapped", "source_mapped", "value", paste("Layer", l, "Head", h), color)
    p
    return(p)
  }
```
#### Process Data
```{r}
pattern_df <- data.frame(layer = layer_pattern, head = head_pattern)
#select the head that are in pattern_df (the whole tuple layer, head)
data <- merge(data, pattern_df, by = c("layer", "head"))
#filter just the relevant positions for both source_position and dest_position
data <- data %>% filter(dest_position == 1 | dest_position==4 | dest_position == 5 | dest_position== 6 | dest_position==8 |  dest_position == 11 | dest_position== 12)
data <- data %>% filter(source_position == 1 | source_position==4 | source_position == 5 | source_position== 6 | source_position==8 |  source_position == 11 | source_position== 12)
#mapped position


unique_positions <- unique(data$dest_position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data$dest_mapped <- unname(position_mapping[as.character(data$dest_position)])
data$source_mapped <- unname(position_mapping[as.character(data$source_position)])
# order the position 
data$dest_mapped <- factor(data$dest_mapped, levels = unique(data$dest_mapped))
data$source_mapped <- factor(data$source_mapped, levels = unique(data$source_mapped))

#select a specific head
data_head <- data %>% filter(layer == 11 & head == 10)

library(ggplot2)

# Reorder factors to have origin at the top and labels for future use
source_mapped_levels <- rev(levels(data_head$source_mapped))
dest_mapped_levels <- levels(data_head$dest_mapped)
scale_y_discrete(limits = source_mapped_levels)
```
#### Plot
```{r}
# plot <- NULL
# for (i in c(1:length(head_pattern))) {
#   head <- head_pattern[i]
#   layer <- layer_pattern[i]
#   if (layer %in% factual_heads_layer & head %in% factual_heads_head){
#     color <-FACTUAL_COLOR
#   }else{
#     color <-  COUNTERFACTUAL_COLOR
#   }
#    plot <- plot + plot_pattern(layer,head, data, color)
# }
# plot + plot_layout(ncol=2, nrow=3)

library(patchwork)

# plots <- list()
# for (i in c(1:length(head_pattern))) {
#   head <- head_pattern[i]
#   layer <- layer_pattern[i]
#   if (layer %in% factual_heads_layer & head %in% factual_heads_head){
#     print(layer)
#     print(head)
#     color <- FACTUAL_COLOR
#   }else{
#     color <-  COUNTERFACTUAL_COLOR
#   }
#   print(color)
#   plots <- c(plots, list(plot_pattern(layer, head, data, color)))
# }

plots <- list()
for (i in c(1:length(head_pattern))) {
  head <- head_pattern[i]
  layer <- layer_pattern[i]
  if ((layer==11 & head==10) || (layer==10 & head==7)){
    print(layer)
    print(head)
    color <- FACTUAL_COLOR
  }else{
    color <-  COUNTERFACTUAL_COLOR
  }
  print(color)
  plots <- c(plots, list(plot_pattern(layer, head, data, color)))
}

plot <- wrap_plots(plots, ncol = 2, nrow = 3)
ggsave(sprintf("tmp/%s_%s_heads_pattern/full_pattern.pdf", model, experiment), plot, width = 80, height = 100, units = "cm")

```

## Boosting Heads
```{r}
data_long <- data.frame(
  model = c("GPT2","GPT2", "Pythia-6.9b",  "Pythia-6.9b"),
  Type = c("Baseline", "Multiplied Attention\nAltered","Baseline", "Multiplied Attention\nAltered" ),
  Percentage = c(4.13,50.29,30.32,49.46)
)
p <- ggplot(data_long, aes(x = model, y = Percentage, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge", color="black", size=1.4) +
  #geom_text(aes(label = label), vjust = -0.5, position = position_dodge(width = 0.9), na.rm = TRUE, size=14) +
  scale_fill_manual(values = c("#ff6361","#003f5c"), labels= c("Baseline", expression(alpha == 5))) +
  labs(x = "",
       y = "% factual answers") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(size = 70, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 70),
    axis.title.y = element_text(size = 75),
    legend.text = element_text(size = 65),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text.align = 0.5, # Center align text relative to the keys
    legend.spacing.x = unit(1.5, "cm")
  ) +
  guides(fill = guide_legend(ncol = 2.5)) # Adjusting the legend
p
ggsave(sprintf("PaperPlot/multiplied_pattern.pdf", model, experiment), p, width = 60, height = 40, units = "cm")
```

# Similarity
```{r}
originaldf <- read_csv(sprintf("%s/evaluate_mechanism_fix_partition.csv", experiment))
#originaldf <- read_csv(sprintf("%s/gpt2_evaluate_mechanism_ss_fixed_partition.csv", experiment))

# Filtering and calculating percentages for self-similarity
df <- originaldf %>% 
  filter(similarity_type == "self-similarity") %>%
  mutate(percentage_true = target_true / (target_true + target_false + other) * 100) %>%
  group_by(model_name, interval) %>%
  summarise(
    percentage_true = mean(percentage_true, na.rm = TRUE),
    n = n(),
    target_true_sum = sum(target_true),
    total = sum(target_true + target_false + other)
  ) %>%
  mutate(ci_lower = binom.confint(target_true_sum, total, methods = "exact")$lower * 100,
         ci_upper = binom.confint(target_true_sum, total, methods = "exact")$upper * 100)


# Calculating percentages for original similarity_type
basedf <- originaldf %>% 
  filter(similarity_type == "original") %>%
  mutate(percentage_true = target_true / (target_true + target_false + other) * 100) %>%
  group_by(model_name) %>%
  summarise(base_percentage = mean(percentage_true, na.rm = TRUE))

# Merging the base percentages with the main DataFrame
df <- df %>% 
  left_join(basedf, by = "model_name")

df <- df %>%
  mutate(model_name = case_when(
    model_name == "gpt2" ~ "GPT2",
    model_name == "gpt2-large" ~ "GPT2-large",
    model_name == "gpt2-medium" ~ "GPT2-medium",
    model_name == "gpt2-xl" ~ "GPT2-xl",
    model_name == "EleutherAI/pythia-6.9b" ~ "Pythia-6.9b",
    TRUE ~ model_name # Keeps other model names unchanged
  ))

# Assuming 'df' is your dataframe and 'interval' is the column to be transformed
percentile_labels <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%", "50-60%", "60-70%", "70-80%", "80-90%", "90-100%")
df$percentile_interval <- factor(df$interval, labels = percentile_labels)

palette <- c("GPT2" = "#003f5c", "GPT2-medium" = "#58508d", "GPT2-large" = "#bc5090", "GPT2-xl" = "#ff6361", "Pythia-6.9b" = "#ffa600")
```

```{r}
p<-ggplot() +
  geom_line(data = df, aes(x = percentile_interval, y = percentage_true, group = model_name, color = model_name), size=1.1) +
  geom_point(data = df, aes(x = percentile_interval, y = percentage_true, group = model_name, color = model_name), size=2.3) +
  geom_line(data = df, aes(x = percentile_interval, y = base_percentage, group = model_name, color = model_name), linetype = "dotted",  size=1.1) +
  scale_color_manual(values = palette) +
  labs(x = "Similarity Score Bins (Percentiles)",
       y = "Percentage of Factual Recalling",
       color = "Model:",
       linetype = "") +
  scale_linetype_manual(values = c("Base Value" = "dotted")) + # Ensure "Base Value" is dotted
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=20),
        axis.text.y= element_text(size=20),
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        axis.title = element_text(size = 23),
        legend.position = "bottom",
        legend.box = "horizontal",
        ) +
        guides(color = guide_legend(nrow = 3, title.position = "top", title.hjust = 0.5),
                                                                                              linetype = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5))
  #save plot
ggsave("PaperPlot/copyVSfact_self_similarity.pdf",p, width = 28, height = 21, units = "cm")
```






# Notepad - not important thing

```{r}
create_heatmap <- function(data, x, y, fill, high_color) {
  n_relevant_position <- length(relevant_position)
  p <- create_heatmap_base(data, x, y, fill) +
    scale_fill_gradient2(low = "white", mid = "white", high = high_color, limits = c(0,31), name = "Logit") +
    #scale_fill_gradient2(low = "black", mid = "white", high = high_color, name = "Logit") +
    #scale_fill_gradient2(low = "black", mid= "white", high = high_color, name = "Logit") +
    
    theme_minimal() +
    #addforce to have all the labels
    scale_x_continuous(breaks = seq(0,n_layers-1 ,1)) +
    scale_y_continuous(breaks = seq(0,n_relevant_position -1,1), labels = relevant_position) +
    scale_y_reverse(breaks = seq(0,n_relevant_position -1,1), labels = relevant_position)+
    labs(x = "Layer", y = "")+
    #fix intenxity of fill
    theme(
      axis.text.x = element_text(size=AXIS_TEXT_SIZE),
      axis.text.y = element_text(size=AXIS_TEXT_SIZE,),
      #remove background grid
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.title.x = element_text(size = AXIS_TITLE_SIZE),
      axis.title.y = element_text(size = AXIS_TITLE_SIZE),
      legend.text = element_text(size = 40),
      legend.title = element_text(size = AXIS_TEXT_SIZE),
      #remove the legend\
      legend.position = "bottom",
      #increase the legend size
      legend.key.size = unit(2.5, "cm"),
      # move the y ticks to the right
    ) 
  return(p)
}
```

### Load and process data
r
```{r}
#print current working directory
data <- read.csv(sprintf("%s/logit_lens/%s/logit_lens_data_logit.csv", experiment, model_folder))
number_of_position <- max(as.numeric(data$position))
data_resid_post <- data %>% filter(grepl("resid_post", component))
data_resid_post$position_name <- positions_name[data_resid_post$position + 1]
relevant_position <- c("Subject", "Relation", "Relation Last")
#filter just the relevant positions
data_resid_post <- data_resid_post %>% filter(position == 1 | position==4 | position == 5)
unique_positions <- unique(data_resid_post$position)
position_mapping <- setNames(seq(0, length(unique_positions) - 1), unique_positions)
# Apply the mapping to create a new column
data_resid_post$mapped_position <- unname(position_mapping[as.character(data_resid_post$position)])
```

### Residual stram heatmaps

```{r}
p_fact <- create_heatmap(data_resid_post, "layer","mapped_position", "mem",  FACTUAL_COLOR)
p_copy <- create_heatmap(data_resid_post, "layer", "mapped_position", "cp",  COUNTERFACTUAL_COLOR)
p_fact
p_copy
ggsave("tmp/heatmap_fact.pdf",p_fact, width = 35, height = 21, units = "cm")


```







